---
title: "1_data_preparation"
format: html
editor: source
---

Load in libraries and data
```{r, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
source(here::here("R/functions.R"))
source(here::here("R/libraries.R"))

load(here::here('data/se_raw.rda'))
load(here::here('data/metadata.rda'))
```

Extract data from SE and convert them to long format
```{r}
#Extract assay and create dataframe
df <- SummarizedExperiment::assay(se_raw)

#Convert to long format
df_long <- df %>% 
    rownames_to_column(var="metabolite") %>% 
    melt(id.var="metabolite") %>% 
    dplyr::rename(sample = variable,
           abundance = value) %>% 
    merge((metadata %>% rownames_to_column("sample")), by="sample") %>%
    dplyr::relocate(metabolite, 1) %>% 
    dplyr::relocate(abundance, .after=id)
```

Calculations (creates: *df_long_ratios*)
```{r}
#Multiply raw arterial and venous samples by haematocrit corrected blood flow
df_long_flow_corrected <- df_long %>% 
    dplyr::mutate(abundance = flow*(abundance * (1-hb)))

#Divide arterial by venous data
df_long_ratios <- df_long_flow_corrected %>% 
    dplyr::group_by(metabolite, id, time) %>% 
    dplyr::summarise(av_dif = abundance[site == "a"] - abundance[site == "v"]) %>%  #Takes ratio and log2 transforms values
    dplyr::mutate(av_dif_log2 = ifelse(av_dif < 0, log2(abs(av_dif))^-1, log(av_dif)))

#Write to excel sheet
writexl::write_xlsx(df_long_ratios, here::here('data/df_long_ratios.xlsx')) 

#Save data
usethis::use_data(df_long_ratios, overwrite = TRUE)

```
Convert to wide format (creates: *se_ratios*)
Now we take our calculated AV ratios and put them back into the standard wide format. We'll save as a SummarizedExperiment for ease
```{r}
#Create new copy of metadata used for merging the "sample" column into the df_long_ratios in next step. Necessary because they will be the column names in wide format data frame. 
metadata2 <- metadata %>% 
    tibble::rownames_to_column("sample") %>% 
    dplyr::select(-c("hb", "flow")) %>% #Filter unnecessary columns
    dplyr::filter(site == 'a') %>%  #Need to filter away half the samples, because in df_long_ratio, there is only one sample per time point (because they're ratios)
    dplyr::mutate(sample2 = sample) %>% 
    tibble::column_to_rownames("sample2")
    
    
#Pivot the long format data frame into wide format
df_wide_ratios <- df_long_ratios %>% 
    dplyr::select(c("metabolite", "id", "time", "av_ratio")) %>%
    merge(metadata2[,c("sample", "id", "time")], by=c("time", "id")) %>%
    dplyr::select(-c("time", "id")) %>% 
    pivot_wider(names_from = sample, values_from = av_ratio) %>% 
    column_to_rownames("metabolite")

#Columns are out of order after pivoting from long to wide. They need to be in order to construct summarized experiment
col_order <- c("s_1", "s_2", "s_3", "s_4", "s_5", "s_11", "s_12", "s_13", "s_14", "s_15", "s_21", "s_22", "s_23", "s_24", "s_25", "s_31", "s_32", "s_33", "s_34", "s_35", "s_41", "s_42", "s_43", "s_44", "s_45", "s_51", "s_52", "s_53", "s_54", "s_55")

df_wide_ratios <- df_wide_ratios[,col_order]

#Create SummarizedExperiments that stores both the raw and the medianscaled df
se_ratios <- PhosphoExperiment(assay = list(ratios = df_wide_ratios), colData=metadata2)

#Save as data file
usethis::use_data(se_ratios, overwrite = TRUE)

```
We can now move on to create our long format summary dataframe (creates: *df_long_ratios_summary*)
```{r}
#Create new long format datasheet and summarize for mean, standard deviation, and median.
df_long_ratios_summary <- df_long_ratios %>% 
    group_by(metabolite, time) %>% 
    summarise(av_ratio_mean = mean(av_ratio),
              av_ratio_sd = sd(av_ratio),
              av_ratio_median = median(av_ratio),
              n = n()
              )

#Write to excel sheet
writexl::write_xlsx(df_long_ratios_summary, here::here('data/df_long_ratios_summary.xlsx')) 

#Save data
usethis::use_data(df_long_ratios_summary, overwrite = TRUE)

```


